{\bf Landslide: A New Race-Finding Tool for 15-410}

### 2015-09-30 landslide-lecture/2015/lecture.tex
update lecture more
improve lecture more

\begin{document}
\renewcommand{\inserttotalframenumber}{36}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

	{\bf Other extreme}: {\em Nothing} is a preemption point
	\begin{itemize}
		\item Good news: Test will finish quickly.
		\item Bad news: Only one execution was checked for bugginess.
		\begin{itemize}
			\item No alternative interleavings explored.
			\item Makes ``no race found'' a weak claim.
		\end{itemize}
	\end{itemize}
	\linegap

	Is there a {\bf ``sweet spot''}?
\end{frame}

@@

\begin{frame}{Challenges}
	Parameters of systematic tests must be kept small.
	\begin{itemize}
		\item Test program length / number of threads
		\item Number of preemption points used
	\end{itemize}
	\linegap

	\begin{center}
		\hspace{-0.25in}\includegraphics[width=0.3\textwidth]{expo.pdf} \\
		\vspace{-1.33in}\hspace{-1.75in}{\rotatebox{90}{\small state space size}} \\
		\vspace{0.15in}\hspace{-0.25in}{\small number of PPs}
	\end{center}
\end{frame}

@@

\begin{frame}{About The Project}
	\textbf{About me}: 5th year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		\item 1st graduate year was 5th year MS, rest in Ph.D. program
		\begin{itemize}
			\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Identifying Bugs}
	Landslide can {\em definitely discover}:
	\begin{itemize}
		\item Assertion failures % Say: note: the better your asserts are..!
		\item Segfaults
		\item Deadlock
		\item Use-after-free / double-free
	\end{itemize}
	\linegap
	Landslide can {\em reasonably suspect}:
	\begin{itemize}
		\item Infinite loop (halting problem)
		\item Data race bugs
	\end{itemize}
\end{frame}

\begin{frame}{What is a Data Race?}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item The threads can be reordered (e.g. no \texttt{cond\_signal()} in between)
	\end{itemize}
	\pause
	\linegap

@@

	Based on experience from past 15-410 student volunteers
	\begin{itemize}
		\item Students worked best with an iterative process
		\item ``Start small, then add more preemption points as time allows''
		\item Landslide now automates this process
	\end{itemize}
	\linegap

@@

\begin{frame}{Iterative Deepening in Landslide}
	Landslide automatically iterates through different configurations of PPs.
	\begin{itemize}
		\item Manages work queue of jobs with different PPs
		\item Each job represents a new state space for Landslide to explore
		\item Prioritizes which jobs are important / likely to finish
		\begin{itemize}
			\item Based on nature of PPs (data races? mutexes?)
			\item Based on estimated completion time
		\end{itemize}
		\item Repeat state space explorations, adding preemption points, until time is exhausted.
	\end{itemize}

@@

\begin{frame}{Test Suite}
	Landslide ships with 6 test cases:
	\begin{itemize}
		\item {\tt thr\_exit\_join}
		\item {\tt paraguay}
		\item {\tt rwlock\_downgrade\_read\_test}
		\item {\tt broadcast\_test} (new)
		\item {\tt paradise\_lost} (new this semester)
		\item {\tt mutex\_test} (new this semester)
	\end{itemize}
\end{frame}

@@

	Landslide will look for data races inside your mutex implementation.
	\begin{itemize}
		\item {\tt xchg}/{\tt cmpxchg}/{\tt xadd}
		\item {\tt mutex->owner = ...}
	\end{itemize}
	\linegap

@@

	Watch your email for more details!
\end{frame}

@@

\begin{frame}{More Evaluation}
	\textbf{Is dynamically adding ``data race'' PPs effective?}
	\begin{itemize}
		\item 3 among preliminary bugs (9-15\%) required data races to expose.
		\item 6 among S'15 student bugs (46\%)
	\end{itemize}
	\linegap

>>>>

\begin{document}
\renewcommand{\inserttotalframenumber}{40}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

	{\bf Other extreme}: {\em Nothing} is a preemption point
	\begin{itemize}
		\item Good news: Test will finish quickly.
		\item Bad news: Only one execution was checked for bugginess.
		\begin{itemize}
			\item No alternative interleavings explored.
			\item Makes ``no race found'' a weak claim.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Preemption Point Example (remember this?)}
	\fbox{
	\begin{tabular}{l}
	\texttt{boolean~want[2]~=~\{~false,~false~\};}\\
	\\
	\begin{tabular}{rll}
		1 & \texttt{want[i]~=~true;} & \\
		  & & \footnotesize (preemption point A) \\
		2 & \texttt{while~(want[j])} & \\
		  & & \footnotesize (preemption point B) \\
		3 & \texttt{~~~~~~~~continue;} & \\
		  & & \footnotesize (preemption point C) \\
		4 & \texttt{//~...critical~section...} & \\
		  & & \footnotesize (preemption point D) \\
		5 & \texttt{want[i]~=~false;} & \\
	\end{tabular}
	\end{tabular}
	}
	\linegap

	{\bf Some preemption points will expose bugs. \\
	Some preemption points don't matter.}

\end{frame}

\begin{frame}{Preemption Point Example (remember this?)}
	\fbox{
	\begin{tabular}{l}
	\texttt{boolean~want[2]~=~\{~false,~false~\};}\\
	\\
	\begin{tabular}{rll}
		1 & \texttt{want[i]~=~true;} & \\
		  & & \footnotesize (preemption point A) \\
		2 & \texttt{while~(want[j])} & \\
		  & & \footnotesize ~ \\
		3 & \texttt{~~~~~~~~continue;} & \\
		  & & \footnotesize ~ \\
		4 & \texttt{//~...critical~section...} & \\
		  & & \footnotesize ~ \\
		5 & \texttt{want[i]~=~false;} & \\
	\end{tabular}
	\end{tabular}
	}
	\linegap

	Here, only preemption point A will trigger a deadlock. {\bf ~}\\
	All other interleavings are benign.{\bf ~}

\end{frame}

@@

%\begin{frame}{Challenges}
%	Parameters of systematic tests must be kept small.
%	\begin{itemize}
%		\item Test program length / number of threads
%		\item Number of preemption points used
%	\end{itemize}
%	\linegap
%
%	\begin{center}
%		\hspace{-0.25in}\includegraphics[width=0.3\textwidth]{expo.pdf} \\
%		\vspace{-1.33in}\hspace{-1.75in}{\rotatebox{90}{\small state space size}} \\
%		\vspace{0.15in}\hspace{-0.25in}{\small number of PPs}
%	\end{center}
%\end{frame}

@@

\begin{frame}{About The Project}
	\textbf{About me}: 5th year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		\item 1st graduate year was 5th year M.S., rest in Ph.D. program
		\begin{itemize}
			\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Identifying Bugs}
	Landslide can {\em definitely discover:}
	\begin{itemize}
		\item Assertion failures % Say: note: the better your asserts are..!
		\item Segfaults
		\item Deadlock
		\item Use-after-free / double-free
	\end{itemize}
	\linegap
	Landslide can {\em reasonably suspect:}
	\begin{itemize}
		\item Infinite loop (halting problem)
		\item Data race bugs
	\end{itemize}
\end{frame}

\begin{frame}{What is a Data Race?}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item The threads can be reordered (e.g., no \texttt{cond\_signal()} in between)
	\end{itemize}
	\pause
	\linegap

@@

	Based on experience from past 15-410 student volunteers
	\begin{itemize}
		%\item Students worked best with an iterative process
		\item ``Start small, then add more preemption points as time allows''
		\item Landslide now automates this process
	\end{itemize}
	\linegap

@@

\begin{frame}{Iterative Deepening in Landslide}
	Landslide automatically iterates through different configurations of PPs.
	\begin{itemize}
		\item Manages work queue of jobs with different PPs
		\item Each job represents a new state space for Landslide to explore
		\item Prioritizes jobs based on estimated completion time
		%\item Prioritizes which jobs are important / likely to finish
		%\begin{itemize}
		%	\item Based on nature of PPs (data races? mutexes?)
		%	\item Based on estimated completion time
		%\end{itemize}
	\end{itemize}
	\linegap

	Repeat state space explorations, adding preemption points, until time is exhausted.

@@

\begin{frame}{Test Suite}
	Landslide ships with 6 approved test cases:
	\linegap

	Standard P2 tests
	\begin{itemize}
		\item {\tt thr\_exit\_join}
		\item {\tt paraguay}
		\item {\tt rwlock\_downgrade\_read\_test}
	\end{itemize}

	New tests
	\begin{itemize}
		\item {\tt broadcast\_test}
		\item {\tt paradise\_lost} (new this semester)
		\item {\tt mutex\_test} (new this semester)
	\end{itemize}
\end{frame}

@@

	Landslide will look for data races inside your mutex implementation.
	\begin{itemize}
		\item {\tt xchg}/{\tt cmpxchg}/{\tt xadd}
		\item {\tt mutex->whose\_turn\_is\_it = ...;}
	\end{itemize}
	\pause
	\linegap

@@

	{\em Really} interested?
	\begin{itemize}
		\item I know of an M.S. thesis topic if you want to do post-OS research.
	\end{itemize}
\end{frame}

@@

\begin{frame}{More Evaluation}
	\textbf{Is dynamically adding ``data race'' PPs effective?}
	\begin{itemize}
		%\item 3 among preliminary bugs (9-15\%) required data races to expose.
		\item 6 among S'15 student bugs (46\%)
		\item 10 among {\tt mutex\_test} bugs (71\%)
	\end{itemize}
	\linegap
